---
title: "Assignment-3"
author: "Sai Vignesh Reddy Cholleti"
date: "03/16/2022"
output:
  html_document: 
    toc: yes
    number_sections: yes
  html_notebook:
    toc: yes
    number_sections: yes
editor_options:
  chunk_output_type: inline
---

<style type="text/css">

div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { /* Header 1 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { /* Header 2 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
</style>

```{r setup, include=FALSE}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
# knitr::opts_knit$set(root.dir = "C:/Users/75CPENG/OneDrive - West Chester University of PA/Documents")
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")
##
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```

# Data Management

## Reshaping and Merging the datasets

<li> Read four datasets from github repository into four objects named *income*, *popul*, *life* and *countries*. </li>
<li> Reshaped the dataset *income* into a longitudinal dataset which contains only three columns country, year and income by using *gather* and *mutate* functions from *dplyr* package. </li>
<li> In the same manner, reshaped the datasets *popul* and *life* into longitudinal data such that the resulting datasets contains only three respective columns. </li>
<li> Then, merged the datasets *life* and *income* into a new dataset *LifeExpIncom* </li>
<li> Merged the above dataset *LifeExpIncom* with dataset *countries* so that the final data set has information about income, life expectancy, and country region.</li>
<li> Finally, Merged the previous resulting data set with dataset *popul* so that the final data set has information about income, life expectancy, population size, and country region, and named it as *InLEPopCtry*</li>
<li> Applied *filter()* function, to select the data for the year 2015 inorder to create the interactive scatter plot for the year 2015. </li>
<br>

```{r}

income <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/income_per_person.csv")
popul <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/population_total.csv")
life <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/life_expectancy_years.csv")
countries <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/countries_total.csv")

income.new <- income %>%
  gather(key = "Year", value = "Income", -geo, na.rm=TRUE) %>%
  mutate(Country = geo, Year = substr(Year,2,5)) %>%
  select(-geo)
#head(income.new)

life.new <- life %>%
  gather(key = "Year", value = "Life", -geo, na.rm=TRUE) %>%
  mutate(Country = geo, Year = substr(Year,2,5)) %>%
  select(-geo)
#head(life.new)

popul.new <- popul %>%
  gather(key = "Year", value = "population", -geo, na.rm=TRUE) %>%
  mutate(Country = geo, Year = substr(Year,2,5)) %>%
  select(-geo)
#head(popul.new)

LifeExpIncom <- merge(income.new, life.new, by = c("Country","Year") , all.x = FALSE)
LifeIncomCoun <- merge(LifeExpIncom, countries, by.x = "Country", by.y = "name", all.x = FALSE)
LifeIncomCoun <- select(LifeIncomCoun, Year, Country, region, Income, Life)
InLEPopCtry <- merge(LifeIncomCoun, popul.new, by = c("Country","Year"), all.x = FALSE)

rm(countries, income, income.new, life, life.new, LifeExpIncom,LifeIncomCoun, popul, popul.new)
df2015 <- filter(InLEPopCtry, Year == 2015)
```

# Interactive Scatter plot

<li> An Interactive scatter plot is created to show the association between the Avg Income and Life Expectancy for the year 2015. </li>
<li> This scatter plot is created by using the *plot_ly* function from *plotly* package </li>
<li> The instructions followed to construct the below interactive scatter plot are </li>
<ol> <li> Set the point size to be proportional to the population size. </li>
<li> Use different colors for different countries. </li>
<li> Choose an appropriate transparency level so that overlapped points can be viewed. </li>
<li> Choose an appropriate color to highlight the point boundary so that partially overlapped points can be easily distinguished. </li>
<li> Included the country name and population size in the hover text. </li> </ol>


```{r}

df2015 <- mutate(df2015, population_m = population/1000000)

fig1 <- plot_ly( data = df2015,
                 x = ~Income,
                 y = ~Life,
                 marker = list(line = list(color = 'black',
                                           width = 0.75)),
                 color = ~Country,
                 size = ~population_m,
                 sizes = c(50,1000),
                 type = "scatter",
                 mode = "markers",
                 alpha = 0.75,
                 hoverinfo = 'text',
                 text = paste("<br><b>Income: $</b>", df2015$Income,
                              "<br><b>Life: </b>", df2015$Life,
                              "<br><b>Population in millions: </b>", df2015$population_m,
                              "<br><b>Country: </b>", df2015$Country ) ) %>%
        layout( showlegend = FALSE,
                plot_bgcolor ='#E2E2E2',
                title = list( text = "<b>Relationship between Avg Income and Life Expectancy in 2015</b>",
                              font = list( size = 20,
                                           color = "darkred") ),
                xaxis = list( title = "<b>Avg Income per person </b>($)",
                              zerolinewidth = 1.5,
                              gridcolor = 'white'),
                yaxis = list( title = "<b>Life Expectancy </b>(years)",
                              zerolinewidth = 1.5,
                              gridcolor = 'white'),
                margin = list( b = 15, l = 25, t = 85, r = 25) )

fig1
```

# Animated Scatter plot

<li> An Animated scatter plot is created to display the change in relationship between Avg income and life expectancy over the years. (considered the years from 1950 to 2018) </li>
<li> This animated plot is created by using aspects/functions of *gganimate* and *ggplot* packages. Also, this animated plot is created by considering the below aspects. </li> 
<ol> <li> Set the point size to be proportional to the population size </li>
<li> Use different colors for different regions. </li>
<li> Choose an appropriate transparency level so that overlapped points can be viewed. </li>
<li> Choose an appropriate color to highlight the point boundary so that partially overlapped points can be easily distinguished.</li> </ol>
<br>

```{r}

InLEPopCtry$Year <- as.numeric(InLEPopCtry$Year)
df50_18 <- filter(InLEPopCtry, Year >= 1950 & Year <= 2018)

fig2 <- ggplot(data = df50_18, mapping = aes( x = Income,
                                              y = Life,
                                              size = population,
                                              color = Country)) +
        geom_point(alpha = 0.5, show.legend = FALSE) +
        scale_color_viridis_d() + 
        scale_size(range = c(2, 12)) +
        scale_x_log10() +
        scale_y_log10() +
        labs( title = list( text = paste("Animated Scatter Plot between Avg Income and Life                                              Expectancy", "from 1950 to 2018", sep = "\n"),
                            color = "darkred"),
              subtitle = 'Year: {frame_time}',
              x = "Avg Income ($)",
              y = "Life expectancy (years)") +
        transition_time(as.integer(Year)) +
        ease_aes('linear')

animate(fig2, renderer = gifski_renderer(), rewind = TRUE)
anim_save("animated_plot.gif", fig2, path = "D:/OneDrive - West Chester University of PA/Documents/Spring'22/STA553-DataViz/git/Images")
```

# Interactive Choropleth map

<li> Firstly, read the data from github and assigned it to a object named *df* </li>
<li> From the above dataset, a sample of 500 observations was randomly selected and assigned it to another object named *df2* </li>
<li> Then using *plot_geo* function from *plotly* package, created the Choropleth map showing a random sample of 500 gas stations in the US. </li>
<li> The information included in the hover: State, County, Address and the ZIP code of the gas station. </li> 
<br>

```{r}

df <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/Gas_stations_POC.csv")

set.seed(71725)
df_500 <- sample_n(df, 500)

g <- list( scope = 'usa',
           projection = list(type = 'albers usa'),
           showland = TRUE,
           landcolor = toRGB("gray95"),
           subunitcolor = toRGB("gray85"),
           countrycolor = toRGB("gray85"),
           countrywidth = 0.5,
           subunitwidth = 0.5 )

l <- list( font = list( family = "sans-serif",
                        size = 12,
                        color = "#000"),
           bgcolor = "#E2E2E2",
           bordercolor = "#FFFFFF",
           borderwidth = 2 )

fig3 <- plot_geo( data = df_500, x = ~xcoord, y = ~ycoord) %>%
        add_markers( text = ~paste("<br><b>State: </b>", STATE,
                                   "<br><b>County: </b>", county,
                                   "<br><b>Address: </b>", ADDRESS,
                                   "<br><b>Zip code: </b>", ZIPnew),
                     color = ~description,
                     colors = c("springgreen3","dodgerblue3"),
                     alpha = 0.5,
                     symbol = "circle", 
                     #marker = list( line = list( color = 'black', width = 0.5)),
                     size = I(30),
                     hoverinfo = "text") %>%
        layout( title = list( text = "<b>500 Random Gas stations in the United States</b>",
                              font = list( size = 20,
                                           color = "darkred" )),
                legend = l,
                margin = list( b = 15, l = 25, t = 85, r = 25),
                geo = g )

fig3
```
<br>
<br>

