---
title: "Assignment-1"
author: "Sai Vignesh Reddy Cholleti"
date: "02/11/2022"
output:
  html_document: 
    toc: yes
    number_sections: yes
  html_notebook:
    toc: yes
    number_sections: yes
editor_options:
  chunk_output_type: inline
---


<style type="text/css">

div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { /* Header 1 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { /* Header 2 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
</style>

```{r setup, include=FALSE}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
# knitr::opts_knit$set(root.dir = "C:/Users/75CPENG/OneDrive - West Chester University of PA/Documents")
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")
##
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```

# The Scatter plot

This R markdown file contains all the code and graphics required for the Assignment-1 of Data Visualization course. Also, the link for this html file can be accessed by <a href="https://htmlpreview.github.io/?https://github.com/VigneshReddy79/STA553-DataViz/blob/main/Assignments/Assignment-1.html">
clicking here!</a>

<li> Downloaded the penguins data set and then uploaded it to my Github data repository. </li>
<li> Found an image of a penguin and uploaded the image to the Github repository. </li>

Read the data set into R from the Github data repository directly.

```{r}
penguins <- read.csv(file = "https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/w02-penguins.csv")
```

## The base Scatter plot between Flipper length and Bill length

The scatter plot between the numerical variables *bill_length_mm* and *flipper_length_mm* from the *penguins* data set with a regression line.

```{r}
plot( penguins$flipper_length_mm, penguins$bill_length_mm,
     xlab = "Flipper length(mm)",
     ylab = "Bill length(mm)",
     main = "Scatter Plot: Flipper length vs Bill length",
     sub = "Scatter plot between flipper length (in mm) and bill length (in mm) with regression line")
model <- lm(penguins$bill_length_mm ~ penguins$flipper_length_mm, data = penguins)
abline(model, lwd=2, col= "red")
```

## The Scatter plot between Flipper length and Bill length with some necessary annotations

The three different species of penguins from the dataset are: 
<ol>
<li> Adelie </li> 
<li> Gentoo </li> 
<li> Chinstrap </li> 
</ol>

The below image differentiates the three different species in our dataset.\
\
<img src= "https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Images/pen_3_cat.jpg">
\

The numerical variables (*bill_length_mm* and *flipper_length_mm*) from the *penguins* data set are used to make a scatter plot and colored the points based on the species.\
Also, modified the scatter plot by following the below instructions:
<ul>
<li> The point size should be proportional to body_mass_g of the corresponding penguin. </li>
<li> insert the image of the penguin you uploaded to the Github repository to the scatter plot. </li>
<li> add a legend to enhance the readability of your plot. </li>
<li> make sure the axes of your plot are appropriately labeled. The title of plot should reflect the content in the plot. </li>
</ul>


```{r}

pen.bill <- penguins$bill_length_mm
pen.flipper <- penguins$flipper_length_mm

plot( pen.flipper, pen.bill, type='n',
     xlab = "Flipper length(mm)",
     ylab = "Bill length(mm)",
     main = "Scatter Plot: Flipper length vs Bill length",
     sub = "Enhanced Scatter plot between flipper length (in mm) and bill length (in mm) from penguins dataset")

Adelie.id = which(penguins$species=="Adelie")
Gentoo.id = which(penguins$species=="Gentoo")
Chinstrap.id = which(penguins$species=="Chinstrap")
size.bm <- penguins$body_mass_g

col.code <- c(alpha("red",0.5), alpha("blue",0.5), alpha("cyan",0.6))

img.url <- "http://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Images/PenguinMeasurements.png"
temp <- tempfile()
download.file(img.url, temp, mode = "wb", cacheOK = FALSE)
img <- readPNG(temp)
raster.img <- as.raster(img)

points(pen.flipper[Adelie.id], pen.bill[Adelie.id], pch = 19, col = col.code[1], cex = size.bm[Adelie.id]/2700)
points(pen.flipper[Gentoo.id], pen.bill[Gentoo.id], pch = 19, col = col.code[2], cex = size.bm[Gentoo.id]/2700)
points(pen.flipper[Chinstrap.id], pen.bill[Chinstrap.id], pch = 19, col = col.code[3], cex = size.bm[Chinstrap.id]/2700)

legend("topleft", c("Adelie","Gentoo","Chinstrap"), col=col.code, pch = 19)
rasterImage(raster.img,218,32,232,47)
text(200, 59, "The points size is proportional to 'body_mass_g'", col = "black", cex = 0.75)

```

Finally,
knitted the RMD to HTML format and uploaded the HTML file to my Github repository.


