---
title: "Final Project"
author: "Sai Vignesh Reddy Cholleti"
date: "05/13/2022"
output:
  html_document: 
    toc: yes
    number_sections: yes
  html_notebook:
    toc: yes
    number_sections: yes
editor_options:
  chunk_output_type: inline
runtime: shiny
---

<style type="text/css">

div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { /* Header 1 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { /* Header 2 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
</style>

```{r setup, include=FALSE}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
# knitr::opts_knit$set(root.dir = "C:/Users/75CPENG/OneDrive - West Chester University of PA/Documents")
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")
##
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(plyr)
library(plotly)
library(shiny)
```

# Final Project


## Data manipulation

A dataset containing the prices and other attributes of almost 54,000 diamonds.
The description of the variables are:
<li> <i>price: </i> price in US dollars (\$326–\$18,823) </li>
<li> <i>carat: </i>weight of the diamond (0.2–5.01) </li>
<li> <i>cut: </i>quality of the cut (Fair, Good, Very Good, Premium, Ideal) </li>
<li> <i>color: </i> diamond colour, from D (best) to J (worst) </li>
<li> <i>clarity: </i> a measurement of how clear the diamond is (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best)) </li>
<li> <i>x: </i> length in mm (0–10.74) </li>
<li> <i>y: </i> width in mm (0–58.9) </li>
<li> <i>z: </i> depth in mm (0–31.8) </li>
<li> <i>depth: </i> total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79) </li>
<li> <i>table: </i> width of top of diamond relative to widest point (43–95) </li>

From the above data, a random sample of 1000 observations (100 observations per variable) are choosen for this project.

```{r}
dmnd <- read.csv("https://raw.githubusercontent.com/VigneshReddy79/STA553-DataViz/main/Datasets/dmnd.csv")

dmnd <- dmnd[,-1]
colnames(dmnd)[c(5,7:10)] <- c("total_depth_percent","price ($)","length (mm)","width (mm)","depth (mm)")


choices = unique(dmnd$cut)
x.choice = names(dmnd)[c(1,5:10)]
y.choice = names(dmnd)[c(1,5:10)]


```

## R code for R Shiny web application

Below is the R code for creating shiny web application. The daiamonds data is analysed by using this web application.

```{r}

ui <- fluidPage(
  titlePanel(h4("Analysis of Diamonds data", 
                align = "center", style = "color:navy", br(),br())),
  sidebarLayout(
    sidebarPanel(
      helpText(
        "This side bar helps to analyze different aspects of data"
      ),
      hr(),
      helpText(
        "This selection corresponds to all tabs"
      ),
      radioButtons(inputId = "cut", 
                   label = "Diamond Cut",
                   choices = c(choices,"All"),
                   inline = FALSE,
                   selected = "All"),
      hr(),
      helpText(
        "This selection corresponds to all tabs"
      ),
      selectInput(inputId="Y",
                  label="Response Variable: Y",
                  choices=y.choice,
                  selected=x.choice[4]),
      helpText(
        "This selection corresponds to tabs 1,4,5 & 6"
      ),
      selectInput(inputId="X",
                  label="Predictor Variable: X",
                  choices=x.choice,
                  selected=x.choice[1]),
      hr(),
      helpText(
        "This selection corresponds to tab 6"
      ),
      numericInput(inputId = "newX", 
                   label = "New Value for Prediction:",
                   value=5, 
                   min=1, 
                   max=18800,
                   step=0.5)
    ),
    mainPanel(
      tabsetPanel(type = "tabs",
                  tabPanel("1.Scatter Plot", plotlyOutput("plotly"), textOutput("plotly_txt")),
                  tabPanel("2.Stats by Clarity", textOutput("stats_clarity_txt"), plotlyOutput("stats_clarity")),
                  tabPanel("3.Stats by Color", textOutput("stats_color_txt"), plotlyOutput("stats_color")),
                  tabPanel("4.Regression Plot", textOutput("plot_txt"), plotOutput("plot")),
                  tabPanel("5.Regression Coefficients", textOutput("table_txt"),tableOutput("table")),
                  tabPanel("6.Prediction", textOutput("pred_txt"), plotOutput("predPlt"))
      )
    )
  )
)


```


```{r}

server <- function(input, output) {
  
  workDat = function(dat){
    if (input$cut == "Ideal") {
      workingData = dmnd[which(dmnd$cut == "Ideal"),]
    } else if (input$cut == "Premium") {
      workingData = dmnd[which(dmnd$cut == "Premium"),]
    } else if (input$cut == "Good") {
      workingData = dmnd[which(dmnd$cut == "Good"),]
    } else if (input$cut == "Very Good") {
      workingData = dmnd[which(dmnd$cut == "Very Good"),]
    } else if (input$cut == "Fair") {
      workingData = dmnd[which(dmnd$cut == "Fair"),]
    } else {
      workingData = dmnd
    }
    workingData
  }
  
  output$plotly <- renderPlotly({
    
    plot_ly( data = workDat(),
             x = ~workDat()[[input$X]],
             y = ~workDat()[[input$Y]],
             color = ~workDat()$cut,
             size = ~workDat()$carat,
             type = "scatter",
             mode = "markers",
             alpha = 0.75,
             hovertemplate = paste('<i><b>',input$X,'<b></i>: %{x}',
                                   '<br><b>',input$Y,'</b>:  %{y}',
                                   '<br><b> carat: ',workDat()$carat,"</b>",
                                   '<br><b>',input$cut,'</b>')
    ) %>% layout( showlegend = FALSE,
                  plot_bgcolor ='#E2E2E2',
                  title = list( text = paste("<b>Relationship between",
                                             input$Y, "and", input$X, "</b>"),
                                font = list( size = 20,
                                             color = "darkred") ),
                  xaxis = list( title = paste("<b>", input$X, "</b>"),
                                gridcolor = 'white'),
                  yaxis = list( title = paste("<b>",input$Y, "</b>"),
                                gridcolor = 'white'),
                  margin = list( b = 15, l = 15, t = 35, r = 15) 
    )
  })
  
  output$plotly_txt <- renderText({
    "\n *Note: The point size is proportional to diamond carat."
  })
  
  output$stats_clarity_txt <- renderText({
    "This tab displays some descriptive statistics of the diamonds data by clarity. \n"
  })
  
  output$stats_clarity <- renderPlotly({
    
    fig1 <- plot_ly( data = workDat(), 
                     labels = ~workDat()$clarity, 
                     type = "pie",
                     name = "By Clarity",
                     textposition = 'inside',
                     textinfo = 'label+percent',
                     insidetextfont = list(color = '#FFFFFF'),
                     marker = list(colors = colors,
                                   line = list(color = '#FFFFFF', width = 1)),
                     domain = list(row = 0, column = 0)
    )
    fig1_ <- subplot( fig1, margin = 0.05)  %>% 
      layout( showlegend = FALSE,
              grid = list( rows=2, columns=2),
              xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
              yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
      )
    fig2 <- plot_ly( type = 'bar',
                     name = paste("Avg", input$Y, "by clarity"),
                     x = ~workDat()$clarity,
                     y = ~workDat()[[input$Y]],
                     marker = list( size = 10,
                                    color = 'blue',
                                    opacity = 0.8),
                     hovertemplate =  paste('<i><b>Avg',input$Y,'<b></i>: %{y}',
                                            '<br><i><b>Color:', workDat()$clarity, '<b></i>'),
                     transforms = list( list( type = 'aggregate',
                                              groups = ~workDat()$clarity,
                                              aggregations = list( list( target = 'y', 
                                                                         func = 'avg', 
                                                                         enabled = T))
                     )
                     )
    )
    fig3 <- plot_ly( data = workDat(),
                     type = "histogram",
                     x = ~workDat()[[input$Y]],
                     name = paste("Distribution of", input$Y),
                     hovertemplate =  paste('<i><b>',input$Y,'<b></i>: %{x}',
                                            '<br><b>Frequency: </b> %{y}')
    )
    fig4 <- plot_ly( type = "box",
                     y = ~workDat()[[input$Y]],
                     x = ~workDat()$clarity,
                     name = "Box plot"
    )
    
    
    annotations = list( list( x = 0.25,
                              y = 1,
                              text = "Frequency of diamonds by clarity",
                              xref = "paper", 
                              yref = "paper",
                              xanchor = "center",
                              yanchor = "bottom",
                              showarrow = FALSE
    ),
    list( x = 0.8,
          y = 1,
          text = paste("Average",input$Y),
          xref = "paper", 
          yref = "paper",
          xanchor = "center",
          yanchor = "bottom",
          showarrow = FALSE
    ),
    list( x = 0.5,  
          y = 0.45,  
          text = paste("Histogram/Box plot showing the distribution of", input$Y), 
          xref = "paper",  
          yref = "paper",  
          xanchor = "center",  
          yanchor = "bottom",  
          showarrow = FALSE
    )
    )
    fig <- subplot( fig1_, fig2, fig3, fig4 ,nrows = 2, margin = 0.05) %>% 
      layout( #title = "Specs with Subplot Title", 
        plot_bgcolor='#e5ecf6', 
        xaxis = list( zerolinecolor = '#ffff', 
                      zerolinewidth = 2,
                      gridcolor = 'ffff'), 
        yaxis = list( zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
        bargap = 0.1,
        annotations = annotations,
        showlegend=FALSE,
        showlegend2=FALSE 
      ) 
    
    
  })
  
  output$stats_color_txt <- renderText({
    "This tab displays some descriptive statistics of the diamonds data by color. \n"
  })
  
  output$stats_color <- renderPlotly({
    
    fig1 <- plot_ly( data = workDat(), 
                     labels = ~workDat()$color, 
                     type = "pie",
                     name = "By Color",
                     textposition = 'inside',
                     textinfo = 'label+percent',
                     insidetextfont = list(color = '#FFFFFF'),
                     marker = list(colors = colors,
                                   line = list(color = '#FFFFFF', width = 1)),
                     domain = list(row = 0, column = 0)
    )
    fig1_ <- subplot( fig1, margin = 0.05)  %>% 
      layout( showlegend = FALSE,
              grid = list( rows=2, columns=2),
              xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
              yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
      )
    fig2 <- plot_ly( type = 'bar',
                     name = paste("Avg", input$Y, "by color"),
                     x = ~workDat()$color,
                     y = ~workDat()[[input$Y]],
                     marker = list( size = 10,
                                    color = 'blue',
                                    opacity = 0.8),
                     hovertemplate =  paste('<i><b>Avg',input$Y,'<b></i>: %{y}',
                                            '<br><i><b>Color:', workDat()$color, '<b></i>'),
                     transforms = list( list( type = 'aggregate',
                                              groups = ~workDat()$color,
                                              aggregations = list( list( target = 'y', 
                                                                         func = 'avg', 
                                                                         enabled = T))
                     )
                     )
    )
    fig3 <- plot_ly( data = workDat(),
                     type = "histogram",
                     x = ~workDat()[[input$Y]],
                     name = paste("Distribution of", input$Y),
                     hovertemplate =  paste('<i><b>',input$Y,'<b></i>: %{x}',
                                            '<br><b>Frequency: </b> %{y}')
    )
    fig4 <- plot_ly( type = "box",
                     y = ~workDat()[[input$Y]],
                     x = ~workDat()$color,
                     name = "Box plot"
    )
    
    
    annotations = list( list( x = 0.25,
                              y = 1,
                              text = "Frequency of diamonds by color",
                              xref = "paper", 
                              yref = "paper",
                              xanchor = "center",
                              yanchor = "bottom",
                              showarrow = FALSE
    ),
    list( x = 0.8,
          y = 1,
          text = paste("Average",input$Y),
          xref = "paper", 
          yref = "paper",
          xanchor = "center",
          yanchor = "bottom",
          showarrow = FALSE
    ),
    list( x = 0.5,  
          y = 0.45,  
          text = paste("Histogram/Box plot showing the distribution of", input$Y), 
          xref = "paper",  
          yref = "paper",  
          xanchor = "center",  
          yanchor = "bottom",  
          showarrow = FALSE
    )
    )
    fig <- subplot( fig1_, fig2, fig3, fig4 ,nrows = 2, margin = 0.05) %>% 
      layout( #title = "Specs with Subplot Title", 
        plot_bgcolor='#e5ecf6', 
        xaxis = list( zerolinecolor = '#ffff', 
                      zerolinewidth = 2,
                      gridcolor = 'ffff'), 
        yaxis = list( zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
        bargap = 0.1,
        annotations = annotations,
        showlegend=FALSE,
        showlegend2=FALSE 
      ) 
    
  })
  
  output$plot_txt <- renderText({
    "This tab displays the scatter plot with regression line. \n"
  })
  
  output$plot <- renderPlot({
    regdata = workDat()
    if (input$cut == "All"){
      m0 = lm(dmnd[[input$Y]]~dmnd[[input$X]])
      plot(dmnd[[input$X]], dmnd[[input$Y]],
           main = paste("Relationship between", input$Y, "and", input$X),
           type = "p",
           pch = 19,
           col = "blue",
           xlab = input$X,
           ylab = input$Y
      )
      abline(m0, lwd = 2, col ="red")
    } else {
      m1 = lm(workDat()[[input$Y]]~workDat()[[input$X]])
      plot(workDat()[[input$X]], workDat()[[input$Y]],
           main = paste("Relationship between", input$Y, "and", input$X),
           type = "p",
           pch = 19,
           col = "blue",
           xlab = input$X,
           ylab = input$Y
      )
      abline(m1, lwd = 2, col = "red")
    }
  })
  
  output$table_txt <- renderText({
    "This tab summarizes the inferential statistics in the regression model. \n "
  })
  
  output$table <- renderTable({
    br()
    br()
    br()
    m0 = lm(workDat()[[input$Y]] ~workDat()[[input$X]])
    regcoef = data.frame(coef(summary(m0)))
    regcoef$Pvalue = regcoef[,names(regcoef)[4]]
    regcoef$Variable = c("Intercept", input$X)
    regcoef[,c(6, 1:3, 5)]
  })
  
  output$pred_txt <- renderText({
    "This tab displays the regression results.\n"
  })
  
  output$predPlt <- renderPlot({
    m3 = lm(workDat()[[input$Y]] ~ workDat()[[input$X]])
    pred.y = coef(m3)[1] + coef(m3)[2]*input$newX
    plot(workDat()[[input$X]], workDat()[[input$Y]], 
         xlab = input$X,
         ylab = input$Y,
         main = paste("Relationship between", input$Y, "and", input$X)
    )
    abline(m3,col = "red",lwd = 1,lty=2)
    points(input$newX, pred.y, pch = 19, col = "red", cex = 2)
  })
}

shinyApp(ui = ui, server = server)


```

## R Shiny Dashboard


<li> This part of Project includes the creation of dashboard by using R Shiny apps. This Dashboard is created by using flex dashboard package within R packages. </li>
<li> The layout employed for this dashboard is multiple column and Input sidebar layout. The sidebar consists some of the options like specifying diamond cut, selecting numerical variable to obtain relationship between them and reporting issues. </li>
<li> The multiple column layout is choosen for the main panel. This Main panel consists all the required plots and figures. Some of the plots like Box plots and Ridge plots are used to display the distribution of response variable in each cut of diamond. </li>
<li>To visit the web application for the R Shiny Dashboard created on shinyapps.io <a href="https://vigneshreddy79.shinyapps.io/Final_project_dashboard/?_ga=2.207065360.261021921.1652461410-1697144491.1643820914" target="_blank"> Click here! <a> </li>


<br>
<br>



